Motional Ground State Cooling Outside the Lamb-Dicke Regime

### 2018-01-30 paper.tex
Update by Kang-Kuen

Our experiment has an overall repetition rate of $2.5$ Hz and
begins by loading a single sodium atom into an optical tweezer from a MOT~\cite{Hutzler2017-LightShifts}.
The tweezer is created by focussing a $700$ nm wavelength laser beam through an NA=$0.55$ objective to an elliptical waist with radii $\{w_{0,x},w_{0,y}\}\approx \{0.70,0.69\}\,\mu$m.
For $45$ mW of optical power, the trapping frequencies along the three axes are
$\{\omega_x,\omega_y,\omega_z\}/2\pi = \{479(4), 492(5), 86(1)\}\ \text{\text{kHz}}$,
where $z$ is the weakly confined axial direction,
and $x$ and $y$ are the tightly confined radial directions.
After each single atom loading attempt, a first image is taken with a $1.5$ ms exposure time
to determine success, which occurs about 50\% of the time due to the loading mechanism~\cite{Schlosser2001}.
During imaging, the atom is cooled via polarization gradient cooling (PGC),
which reduces the temperature of the single atom to $80$ $\mu$K,
corresponding to mean motional states along the three trap axes of
$\{\bar n_x, \bar n_y, \bar n_z\}=\{3.5(3),\, 3.2(3),\, 20(4)\}$, and therefore initial 3D ground state probability of $P_0=0.4$\%.
The atom is then initialized into
the $|F=2, m_F=2\rangle$ stretched state via optical pumping (OP).
Each experimental cycle ends with a second imaging sequence that measures the hyperfine state of the atom as $|F=2\rangle$ or $|F=1\rangle$~\footnote{The final hyperfine state sensitive detection employs a strong beam that is resonant with the $F=2$ to $F'=3$ cycling transition, to heat the $F=2$ population out of the trap. Therefore, any atoms which survive this heat out procedure are interpreted as having been in the $F=1$ state.}. The data analysis includes only those experiments where the first image reveals successful atom loading.

@@

We thank J. D. Hood for discussions.
This work is supported by the Arnold and Mabel Beckman Foundation, the AFOSR Young Investigator Program, the NSF through the CUA,
and the Alfred P. Sloan Foundation.

>>>>

Our experiment has an overall repetition rate of $2.5$ Hz and
begins by loading a single sodium atom into an optical tweezer from a MOT~\cite{Hutzler2017-LightShifts}.
The tweezer is created by focussing a $700$ nm wavelength laser beam through an NA=$0.55$ objective to a waist with slight elliptical radii $\{w_{0,x},w_{0,y}\}\approx \{0.70,0.69\}\,\mu$m.
For $45$ mW of optical power, the trapping frequencies along the three axes are
$\{\omega_x,\omega_y,\omega_z\}/2\pi = \{479(4), 492(5), 86(1)\}\ \text{\text{kHz}}$,
where $z$ is the weakly confined axial direction,
and $x$ and $y$ are the tightly confined radial directions.
After each single atom loading attempt, a first image is taken with a $1.5$ ms exposure time
to determine success, which occurs about 50\% of the time due to the loading mechanism~\cite{Schlosser2001}.
During imaging, the atom is cooled via polarization gradient cooling (PGC),
which reduces the temperature of the single atom to $80$ $\mu$K,
corresponding to mean motional states along the three trap axes of
$\{\bar n_x, \bar n_y, \bar n_z\}=\{3.5(3),\, 3.2(3),\, 20(4)\}$, and therefore initial 3D ground state probability of $P_0=0.4$\%.
The atom is then initialized into
the $|F=2, m_F=2\rangle$ stretched state via optical pumping (OP).
Each experimental cycle ends with a second imaging sequence that measures the hyperfine state of the atom as $|F=2\rangle$ or $|F=1\rangle$~\footnote{The final hyperfine state sensitive detection employs a strong beam that is resonant with the $F=2$ to $F'=3$ cycling transition, to heat the $F=2$ population out of the trap. Therefore, any atoms which survive this heat out procedure are interpreted as having been in the $F=1$ state.}. The data analysis includes only those experiments where the first image reveals successful atom loading.

@@

This work is supported by the Arnold and Mabel Beckman Foundation, the AFOSR Young Investigator Program, the NSF through the CUA,
and the Alfred P. Sloan Foundation.
